---
title: "Linear modeling"
output: html_notebook
---

In summary, the behavioral data clearly demonstrates the relationship between raection time and correct responses, but only offers weak evidence that meth users perform poorly compared to non-meth users.

```{r, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
library(R.matlab)
library(data.table)
library(ggplot2)
library('tidyr')

library(corrplot)
library(Matrix)
require(lme4)
nonResponseTimeAsNA<-TRUE
source("rl_behav_analysis_learning_setup.R")
```


```{r, message=FALSE, warning=FALSE, include=FALSE}
#put this in the right format
rl.all.subjects.dt<-data.table(rl.all.subjects.list)

#some helper functions.
fixedef.se<-function(fit){
  return(summary(fit)$coefficients[,"Std. Error"])
}

lm.beta.lmer <- function(mod) {
  b <- fixef(mod)[-1]
  sd.x <- apply(getME(mod,"X")[,-1],2,sd)
  sd.y <- sd(getME(mod,"y"))
  b*sd.x/sd.y
}

rl.all.subjects.list$reaction_time_adj<-rl.all.subjects.list$reaction_time
rl.all.subjects.list[is.na(reaction_time),reaction_time_adj:=1]
#table(is.na(rl.all.subjects.list$reaction_time))
#table(is.na(rl.all.subjects.list$reaction_time_adj))
rl.all.subjects.list.clean<-rl.all.subjects.list[score!=0]

```

After eliminating `r rl.all.subjects.list[score==0,.N]` non-response trials across the subjects, `r rl.all.subjects.list[score!=0,.N]` trials remain. 

```{r}


#what does reaction time look like in a distribution?

ggplot(rl.all.subjects.list.clean,aes(reaction_time,color=correct))+geom_density()+#facet_grid(SexRisk~.,margins=TRUE)
  scale_x_continuous(breaks=seq(0.2,1.2,0.2))+
  labs(title="Reaction time density by outcome")
# ggplot(rl.all.subjects.list[score!=0],aes(reaction_time_adj,color=correct))+geom_density()+facet_grid(SexRisk~.,margins=TRUE)+
#   scale_x_continuous(breaks=seq(0.2,1.2,0.2))+
#   labs(title="Reaction time density by outcome (color) and sexrisk (axis)")


```

Correct trials clearly are somewhat slower than incorrect trials, indicating that waiting, on average, paid off for participants.

There is some sign of a bimodal distribution for the incorrect trials. A small number of trials may be clustered around 0.4 s, and then a second set clustered around 0.62 s, just a few milliseconds below the peak of the correct trials at 0.65 s. \footnote{Should posit a bimodal distribution in a simple stan model and see if it does better than a unimodal model.}


I applied a logistic regression model to predict correct responses using reaction time, the position of each image in the segment, and the groups, presented as separate variables (MethUse and SexRisk), with subject ID as a random effect. \footnote{Really, runID also needs to be in here as a random effect, and actually, we need runid to be a random effect from 1 to 4}
```{r cache=TRUE}

mA2c<-glmer(correct~ 
                 #reaction_time+
                  presentation_n_in_segment+ #trial-level
                 #MethUse+
              SexRisk+(1|subid), #subject-level
               data = rl.all.subjects.list.clean,
            family = binomial, control=glmerControl(optimizer="Nelder_Mead"))
summary(mA2c)

mA2c.nomethuse.rt0<-glmer(correct~ 
                 reaction_time+presentation_n_in_segment+ #trial-level
                 SexRisk+(1|subid), #subject-level
               data = rl.all.subjects.list.clean,
               family = binomial, control=glmerControl(optimizer="Nelder_Mead"))
summary(mA2c.nomethuse.rt0)
```

```{r}
mA2c.rt0<-glmer(correct~ 
                 reaction_time+presentation_n_in_segment+ #trial-level
                 MethUse+SexRisk+(1|subid), #subject-level
               data = rl.all.subjects.list.clean,family = binomial, control=glmerControl(optimizer="Nelder_Mead"))
summary(mA2c.rt0)


lm.beta.lmer(mA2c.rt0)

```

The model clearly links reaction time with correct responses. We can calculate standardized $\beta$ values. The standardized beta for reaction time is 0.22 and the standardized beta for SexRisk, which also appeared significant, is 0.10.

```{r}

anova(mA2c,mA2c.nomethuse.rt0,mA2c.rt0)

```

An ANOVA comparison of the models offers unambiguous evidence with a strong difference in both AIC and BIC that adding reaction to the model improves its prediction of a correct response. However, the addition of MethUse as a predictor only improves AIC slightly, by 4 points, and the BIC, which more strongly peanlizes model complexity, suggests that the model including MethUse is not as powerful.

```{r}
# does eliminating reaction time change the prediction?
# in other words, does reaction time carry useful information which might be moderating the rleationship between corrct and MethUse/SexRisk; is reaction time a NEGATIVE CONFOUNDER?

mA2c.noreactiontime<-glmer(correct~ 
                             presentation_n_in_segment+ #trial-level
                             MethUse+SexRisk+(1|subid), #subject-level
                           data = rl.all.subjects.list.clean,family = binomial, control=glmerControl(optimizer="Nelder_Mead"))
summary(mA2c.noreactiontime)

#looking at this - the regression coefficients for MethUse and SexRisk appear to disappear
#Reaction time may be a negative confounder.
#https://www.researchgate.net/post/Why_in_regression_analysis_does_the_inclusion_of_a_new_variable_makes_other_variables_that_previously_were_not_statistically_significant

mA2c.rt.interact<-glmer(correct~ 
                 reaction_time+presentation_n_in_segment+ #trial-level
                 MethUse*reaction_time+SexRisk*reaction_time+(1|subid), #subject-level
               data = rl.all.subjects.list.clean,family = binomial, control=glmerControl(optimizer="Nelder_Mead"))
summary(mA2c.rt.interact)
#quite interesting: allowing an interaction term between reaction_time and MethUse,
# - MethUse becomes more significant
# - Impact of SexRisk on performance disappears
# - The interaction of RT with MethUse and SexRisk are not themselves significant.

rt.A<-lmer(reaction_time~ 
                          +presentation_n_in_segment+ #trial-level
                          MethUse+(1|subid), #subject-level
                        data = rl.all.subjects.list.clean)
summary(rt.A)
#MethUse doesn't predict reaction time....
rt.A1<-lmer(reaction_time~ correct*MethUse
             +presentation_n_in_segment+ #trial-level
             MethUse+(1|subid), #subject-level
           data = rl.all.subjects.list.clean)
summary(rt.A1)

#But MethUse interacting with correct might be a predictor; i.e.,...
rt.A1.cor<-lmer(reaction_time~ MethUse
            +presentation_n_in_segment+ #trial-level
              (1|subid), #subject-level
            data = rl.all.subjects.list.clean[correct==TRUE])
summary(rt.A1.cor)

rt.A1.ncor<-lmer(reaction_time~ MethUse
                +presentation_n_in_segment+ #trial-level
                  (1|subid), #subject-level
                data = rl.all.subjects.list.clean[correct==FALSE])
summary(rt.A1.ncor)
#Nope. Neither in Correct nor incorrect trials alone is MethUse a predictor of reaction time.

#SO, it is clear here that:
# - reaction time is STILL a predictor of performanace
```



```{r}
anova(mA2c,mA2c.noreactiontime,mA2c.rt.interact)
```


```{r}
#DO A LINEAR MODEL PREDICTING NON-RESPONSE RATES FROM METH USE AND OTHER RELEVANT VARIABLES.
#THIS MIGHT IMPACT ON THE WAY WE TREAT THE DATA.


#Here, where we include interactions of reaction time being less than 0.1 or more than 0.1,
#we can see that when raection time is less than 0.1, nothing is particularly predictive of score.
#however, when reaction time is greater than 0.1, MethUse, presentation_n, etc, are predictive.
#digging a bit deeper, performance these trials with low reaction times is worse than chance:

table(rl.all.subjects.list[,.(reaction_time>0.1,correct)])

#This is 98% INCORRECT! Something is off here. No one's performance could be this bad by chance.
#what's going on?
#- this is 

ggplot(rl.all.subjects.list.clean,aes(reaction_time,color=correct))+geom_density()+#facet_grid(SexRisk~.,margins=TRUE)
  coord_cartesian(xlim=c(0,0.2))
  labs(title="Reaction time density by outcome")

hist(rl.all.subjects.list[reaction_time<0.2,reaction_time],breaks = 100)
#ahh. So there is a mistake in the data.
#We have 8000 trials where the response is recorded as incorrect but their reaction time is 0.
table(rl.all.subjects.list$correct)
#need to work out why these were coded as incorrect when they were really non-responses!
```


