---
title: "drift rate example"
author: "Ben Smith"
date: "6/4/2018"
output: pdf_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(LaplacesDemon)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```


```{r ev_to_function}

expected_value <- seq(-1,1,0.01)
f_ev_1<-logit((expected_value+3)/4)
f_ev_2<-expected_value+1
mappings<-data.frame("Expected Value"=expected_value,"f_1"=f_ev_1,"f_2"=f_ev_2) %>% tidyr::gather("Function","Value",f_1:f_2)

ggplot(mappings,aes(x=Expected.Value,y=Value,group=Function,color=Function))+geom_line()

```

Applying a model, we would see expected value proceeding along a curve as the subject learns. With a learning rate of alpha=0.2, and steady learning, this would look like:

```{r ev_over_time}
alpha<-0.2
ev<-0
iter<-20
ev_rec<-rep(NA,iter)
for (i in 1:iter){
  ev_rec[i]<-ev
  pe<-1-ev
  ev<-ev+pe*alpha
  
}
mappings2<-data.frame("Expected Value"=ev_rec,"learning_iteration"=1:iter)
ggplot(mappings2,aes(x=learning_iteration,y=Expected.Value))+geom_line()

```

Thus, the two functions would proceed as the subject learns as:


```{r ev_and_functions_over_time}
mappings2$f_1<-logit((mappings2$Expected.Value+3)/4)
mappings2$f_2<-mappings2$Expected.Value+1
mappings2_ggplot<-tidyr::gather(mappings2,"Function","Value",c(Expected.Value,f_1,f_2))
ggplot(mappings2_ggplot,aes(x=learning_iteration,y=Value,group=Function,color=Function))+geom_line()+labs(y="Drift rate")

```

Thus, as long as the subject steadily learns, the drift rate will increase linearly with learning in the first model, whereas it reaches an asymptote with the other models. A linearly, infinitely increasing drift rate seems not plausible; however, consider that this parameter itslef operates according to a function....

