---
title: "Exploring Revision 5a MCMC and vb"
output:
  pdf_document: 
    keep_tex: yes
  html_notebook: default
---

Revision 5a has the following features:

 - includes one group of subjects
 - Has a level for multiple runs as random effects
 - Runs can be either reward, punishment, or unspecified. Each subject has an individual parameter specifying difference between reward and punishment runs, and these are drawn from a group-level distribution of runs.
 
The variational bayes estimates for Revision 5 looked good. So I ran Revision 5 on MCMC:

 - on each of the three groups
 - Using both variational bayes and MCMC
 - three times so that we could look at consistency across different runs.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, cache = TRUE)
knitr::opts_knit$set(root.dir="../")
```

```{r setup2, include=FALSE}
source("../util/apply_local_settings.R")
apply_local_settings()
knitr::opts_chunk$set(cache.path = paste0(localsettings$data.dir,"knitrcache"))
source("nate_files/fitGroupsV3Onegroup.R")
source("data_summarize.R")
library(data.table)
library(ggplot2)
source("visualization/geom_hdi.R")
```

##Variational Bayes: posterior comparison

Though we have already done a variational Bayes estimate (see rev5_exploration-vb.Rmd), here, these are presented side-by-side with MCMC analyses for comparison.
```{r }
source("du_model_rev5a-mcmc1timerun.R")
```


```{r getDataIntoTable, echo=FALSE}
#arrange all the data into a single data table.
model.summary.all<-NULL

#iterations
miters<-unlist(lapply(model.summaries,function(m){
  return(length(m$summaryObj$iter))
}))
for(ms.i in 1:length(model.summaries)){
  #ms.i=2
  ms<-model.summaries[[ms.i]]
  ms.summaryObj<-ms$summaryObj
  ms.summaryObj$TestId<-ms.i
  ms.summaryObj$Group<-ms$g
  ms.summaryObj$ModelName<-ms$m
  ms.summaryObj$AnalysisRepetition<-ms$t
  ms.summaryObj$EstimationMethod<-ms$EstimationMethod
  #because when we ran this, we hadn't explicitly recorded estimation methods; 
  #but these are distinguishable by the number of iterations.
  if(is.null(model.summary.all)){
    model.summary.all<-ms.summaryObj
  }else{
    model.summary.all<-rbind(model.summary.all,ms.summaryObj,fill=TRUE)
  }
}
model.summary.all$EstimationMethod<-factor(model.summary.all$EstimationMethod)

#table(model.summary.all$TestId,model.summary.all$EstimationMethod)
```



```{r, echo=FALSE}


ggplot(model.summary.all[Statistic=="mu" & EstimationMethod=="variationalbayes"],aes(x=Value ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)
                                                               ))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Group~Parameter,scales = "free")+
    
    labs(title=paste0("mu statistic (all rounds), variable number of runs model"))



ggplot(model.summary.all[Statistic=="rew_mu" & EstimationMethod=="variationalbayes"],aes(x=Value ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)
                                                               ))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Group~Parameter,scales = "free")+
    
    labs(title=paste0("mu statistic (all rounds), Reward Runs Only"))

ggplot(model.summary.all[Statistic=="pun_mu" & EstimationMethod=="variationalbayes"],aes(x=Value ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)
                                                               ))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Group~Parameter,scales = "free")+
    
    labs(title=paste0("mu statistic (all rounds), Punishment Runs Only"))


```


These all look plausible. I think for the first time, I've got a model that estimates repeated runs and reward and punishment within the same model, and yields separate estimates for reward and punishment by using a parameter that separates the two.

While they all look *plausible*, estimates vary widely, indicating that we need to look at MCMC in order to calculate the outcomes for these values. Fortunately, this current model isn't unwieldy and I should be able to run through MCMC analyses in a tractable period of time.

## Differences

I want to take a look at differences:

 - between reward and punishment runs
 - between Group 2 and Group 3
 - An interaction of those differences.

```{r, echo=FALSE}

#to get difference between reward and punishment runs, need to get rew_* and pun_*, reshape to put them on the same col, then calculate difference.
model.summary.all.rewpun<-model.summary.all[Statistic %in% c("rew_mu","pun_mu")] %>% tidyr::spread(Statistic,Value)

model.summary.all.rewpun[,rew_minus_pun_mu:=rew_mu-pun_mu]

ggplot(model.summary.all.rewpun[EstimationMethod=="variationalbayes"],aes(x=rew_minus_pun_mu ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Group~Parameter,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), alpha, Reward Minus Punishment"))

#graph reward vs. punishment

ggplot(model.summary.all.rewpun[Parameter=="alpha" & EstimationMethod=="variationalbayes"],aes(x=rew_mu,y=pun_mu ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)))+
    geom_point(alpha=0.1)+
    facet_grid(.~Group,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), alpha, Reward and Punishment"))

ggplot(model.summary.all.rewpun[Parameter=="beta" & EstimationMethod=="variationalbayes"],aes(x=rew_mu,y=pun_mu ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)))+
    geom_point(alpha=0.1)+
    facet_grid(.~Group,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), beta, Reward and Punishment"))

```



```{r}

#group 2 compared to group 3.
model.summary.all.notestid<-model.summary.all[,TestId:=NULL] 

model.summary.all.groupcompare<- tidyr::spread(model.summary.all.notestid,Group,Value,sep="")
model.summary.all.groupcompare$Group3_minus_Group2<-
  model.summary.all.groupcompare$Group3-model.summary.all.groupcompare$Group2

ggplot(model.summary.all.groupcompare[EstimationMethod=="variationalbayes"],aes(x=Group3_minus_Group2 ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)
                                                               ))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Statistic~Parameter,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), Group 3 Minus Group 2"))


```

These all look like plausible estimates of group-level parameters.

## Exploring the structure of the parameters

Can we take a peak at how .....


# MCMC results

Let's take a look at the same distributions using MCMC.


```{r, echo=FALSE,cache=FALSE}


ggplot(model.summary.all[Statistic=="mu" & EstimationMethod=="MCMC"],aes(x=Value ,fill=factor(Group),color=factor(Group)
                                                               ))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(.~Parameter,scales = "free")+
    
    labs(title=paste0("mu statistic (all rounds), variable number of runs model"))




ggplot(model.summary.all[Statistic %in% c("rew_mu","pun_mu") & EstimationMethod=="MCMC"],aes(x=Value ,fill=factor(Group),color=factor(Group)
                                                               ))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Statistic~Parameter,scales = "free")+
    
    labs(title=paste0("mu statistic (all rounds), Reward and Punishment Runs"))


```



```{r, echo=FALSE}

#to get difference between reward and punishment runs, need to get rew_* and pun_*, reshape to put them on the same col, then calculate difference.
ggplot(model.summary.all.rewpun[EstimationMethod=="MCMC"],aes(x=rew_minus_pun_mu ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Group~Parameter,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), alpha, Reward Minus Punishment"))

#graph reward vs. punishment

ggplot(model.summary.all.rewpun[Parameter=="alpha" & EstimationMethod=="MCMC"],aes(x=rew_mu,y=pun_mu ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)))+
    geom_point(alpha=0.1)+
    facet_grid(.~Group,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), alpha, Reward and Punishment"))

ggplot(model.summary.all.rewpun[Parameter=="beta" & EstimationMethod=="MCMC"],aes(x=rew_mu,y=pun_mu ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)))+
    geom_point(alpha=0.1)+
    facet_grid(.~Group,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), beta, Reward and Punishment"))

```

Need to check the scale these reward and punishment parameters are on. Are they transformed into the actual space there implemented within?


```{r}

#group 2 compared to group 3.
model.summary.all.notestid<-model.summary.all[,TestId:=NULL] 

model.summary.all.groupcompare<- tidyr::spread(model.summary.all.notestid,Group,Value,sep="")
model.summary.all.groupcompare$Group3_minus_Group2<-
  model.summary.all.groupcompare$Group3-model.summary.all.groupcompare$Group2


ggplot(model.summary.all.groupcompare[EstimationMethod=="MCMC"],aes(x=Group3_minus_Group2 ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)
                                                               ))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Statistic~Parameter,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), Group 3 Minus Group 2"))
head(model.summary.all.groupcompare)

model.summary.all.g3g2compare.bypar<- 
  tidyr::spread(
    model.summary.all.groupcompare[,.(iter,Run,Statistic,Parameter,ModelName,AnalysisRepetition,
                                   EstimationMethod,Group3_minus_Group2)],
    Parameter,
    Group3_minus_Group2)

ggplot(model.summary.all.g3g2compare.bypar[
  EstimationMethod=="MCMC" & 
  Statistic %in% c("mu","rew_mu","pun_mu")
                                              ],aes(x=alpha,y=beta 
                                                    #,fill=factor(Statistic),color=factor(Statistic)
                                                    ))+
    geom_point(alpha=0.1)+
    facet_grid(~Statistic,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), Group 3 Minus Group 2"))


```




```{r}

#group 1 compared to group 2
model.summary.all.groupcompare$Group2_minus_Group1<-
  model.summary.all.groupcompare$Group2-model.summary.all.groupcompare$Group1

ggplot(model.summary.all.groupcompare[EstimationMethod=="MCMC"],aes(x=Group2_minus_Group1 ,fill=factor(AnalysisRepetition),color=factor(AnalysisRepetition)
                                                               ))+
    geom_freqpoly(alpha=0.9,binwidth=0.001)+
     geom_hdi(size=2, lineend = "round",alpha=0.5,credible_mass=0.95)+
    facet_grid(Statistic~Parameter,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), Group 2 Minus Group 1"))



model.summary.all.g2g1compare.bypar<- 
  tidyr::spread(
    model.summary.all.groupcompare[,.(iter,Run,Statistic,Parameter,ModelName,AnalysisRepetition,
                                   EstimationMethod,Group2_minus_Group1)],
    Parameter,
    Group2_minus_Group1)

ggplot(model.summary.all.g2g1compare.bypar[
  EstimationMethod=="MCMC" & 
  Statistic %in% c("mu","rew_mu","pun_mu")
                                              ],aes(x=alpha,y=beta))+
    geom_point(alpha=0.02)+
    facet_grid(~Statistic,scales = "free")+
    labs(title=paste0("mu statistic (all rounds), Group 2 Minus Group 1"))


```

