---
title: "hierarchical_summary"
author: "Ben Smith"
date: "7/15/2018"
output: 
  pdf_document: 
    keep_tex: yes
---

## A two-level hierarchical model of behavioral analysis of all subjects across the task.

This dataset was comprised of data from just one reward run, and was a hierarchical model across all three groups.

In this dataset we found that 
I analyzed a number of subjects and found that across the group, meth users had lower learning rates than non-meth users.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	tidy = TRUE
)
knitr::opts_knit$set(root.dir="../")

```

```{r setup2}

source("../util/apply_local_settings.R")
apply_local_settings()
knitr::opts_chunk$set(cache.path = paste0(localsettings$data.dir, "knitrcache"), warning = FALSE)
source("nate_files/fitGroupsV3Onegroup.R")
source("data_summarize.R")
source("data_summarize_lbarlmodel.R")
library(data.table)
library(ggplot2)
mainDataDir=localsettings$data.dir
```


```{r setup_flat_behavioral_analysis}

library(ggplot2)

source("../joint_msm_combined/bjs_misc_utils.R")
version="h_m4"
save.name=paste0("main_", version)
source('de_mcmc/functions.R')
source('de_mcmc/main_m1_setup.R')
source('de_mcmc/functions_joint_v2.R')
source(paste0('de_mcmc/functions_',version,'.R'))

source("de_mcmc/functions_h_m3.R")
source("visualization/geom_hdi.R")
load(paste(save.dir,"output_h_m420180421T172527",".RData",sep=""))
#load.image(paste(save.name,"run.ts",".RData",sep=""))
setwd(mainDir)

start=2
start.weights=2

pdf(paste(save.dir,save.name,".pdf",sep=""),10,5)
par(mfrow=c(1,2),ask=FALSE)
setwd(mainDir)
source("de_mcmc/fig_base5.R")
hpar.names<-paste0(dimnames(phi)[[2]],"_",rep(dimnames(phi)[[3]],times=1,each=6))
phi<-matrix(phi,nrow = dim(phi)[1],ncol=prod(dim(phi)[2:3]))
n.hpars<-18
#fig_base(env = environment(),plot.phi=TRUE,plot.lower = TRUE,ask=FALSE)
fig_base( env = environment(),plot.phi=TRUE,plot.lower = FALSE,ask=FALSE,
          par.functions=rep(c(f_alpha_s_tr,f_thresh_s_tr,f_tau_s_tr),6))#list one function for each param. the second set are for the sigmas.
dev.off()

#now let's do some more plots.
additional_output<-paste(save.dir,save.name,"_extended",sep="")
dir.create(additional_output)


#cool
```

This is derived from raw_data_all_runs_flat_v2.R.

```{r flat_behavioral_analysis}


start=2
start.weights=2

#now let's look at the relationships between each of the phi values.

source("plot_library.R")
res<-hyper_param(
  param.names = par.names.l2,
  group.names = l2.groups.list,
  tnmc = 1001,
  n.chains = n.chains)

#dim(res)
par(mfrow=c(3,3),ask=FALSE)
sr_diff<-res[,,,"Risky No Meth"]-res[,,,"Safe No Meth"]
hist(sr_diff[,,"alpha_mu"],breaks=100,main="Risky-Safe No Meth alpha_mu")
hist(sr_diff[,,"thresh_mu"],breaks=100,main="Risky-Safe No Meth thresh_mu")
hist(sr_diff[,,"tau_mu"],breaks=100,main="Risky-Safe No Meth tau_mu")
HDIofMCMC(sr_diff[,,"tau_mu"])

meth_diff<-res[,,,"Risky Meth"]-res[,,,"Risky No Meth"]
table(meth_diff[,,"alpha_mu"]<0)/prod(dim(meth_diff[,,"alpha_mu"]))
hist(meth_diff[,,"alpha_mu"],breaks=100,main="Meth-NoMeth (sexually risky) alpha_mu")
HDIofMCMC(meth_diff[,,"alpha_mu"])
hist(meth_diff[,,"thresh_mu"],breaks=100,main="Meth-NoMeth (sexually risky) thresh_mu")
HDIofMCMC(meth_diff[,,"thresh_mu"])
hist(meth_diff[,,"tau_mu"],breaks=100,main="Meth-NoMeth (sexually risky) tau_mu")
HDIofMCMC(meth_diff[,,"tau_mu"])

sr_meth_diff<-res[,,,"Risky Meth"]-res[,,,"Safe No Meth"]
hist(sr_meth_diff[,,"alpha_mu"],breaks=100,main="Risky Meth - Safe NoMeth alpha_mu")
HDIofMCMC(sr_meth_diff[,,"alpha_mu"])
hist(sr_meth_diff[,,"thresh_mu"],breaks=100,main="Risky Meth - Safe NoMeth thresh_mu")
hist(sr_meth_diff[,,"tau_mu"],breaks=100,main="Risky Meth - Safe NoMeth tau_mu")
dim(res[,,,"Risky No Meth"])
```

```{r}
res3<-array(res,dim=c(prod(dim(res)[1:2]),dim(res)[3],dim(res)[4]),
            dimnames=dimnames(res)[2:4])

# res3[2,,"Risky Meth"]
# res[2,1,,"Risky Meth"]
# res3[2,"alpha_mu",]
# res[2,1,"alpha_mu",]
alpha_mu<-data.table(res3[,"alpha_mu",])
alpha_mu_graph<-gather(alpha_mu,key = "Group","alpha_mu")

group_plot=ggplot(alpha_mu_graph,
       aes(x=alpha_mu,color=Group))+geom_density(size=2)+geom_hdi(alpha=0.8,size=2)+
  labs(title="standardized alpha_mu posterior estimates by group")

group_plot

```

```{r}
#print a table of the 95% HDIs
#columns are parameters; rows are group comparisons
hdi.table<-NULL
group.comparisons<-list("sr_diff"=sr_diff,"meth_diff"=meth_diff,"sr_meth_diff"=sr_meth_diff)

HDI_format<-function(hdi_pair,dp=NA){
  if(is.na(dp)){
    return(paste0("[",hdi_pair[1],", ",hdi_pair[2],"]"))
  }else{
    return(paste0("[",format(hdi_pair[1],digits=dp,nsmall=dp,scientific=-2),", ",format(hdi_pair[2],digits=dp,nsmall=dp,scientific=-2),"]"))
  }
  
}

for (gc_name in names(group.comparisons)){
  gc=group.comparisons[[gc_name]]
  hdi.table.r<-data.frame("alpha_mu" = HDI_format(HDIofMCMC(gc[,,"alpha_mu"]),3),
                          "thresh_mu" = HDI_format(HDIofMCMC(gc[,,"thresh_mu"]),3),
                          "tau_mu" = HDI_format(HDIofMCMC(gc[,,"tau_mu"]),3)
                          )
  rownames(hdi.table.r)[1]<-gc_name
  if (is.null(hdi.table)){
    hdi.table <- hdi.table.r
  }else{
    hdi.table<-rbind(hdi.table,hdi.table.r)
  }
  
}
```

## Parameters

Here is the HDI for each of the parameters for this model. 

The learning rate difference between meth-using subjects and non-meth-using subjects was significant.

```{r flat_behavioral_analysis_table}

knitr::kable(hdi.table)
 
```

